---
title: "Assignment 2: Coding Basics"
author: "April Qian"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics. 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
2. Work through the steps, **creating code and output** that fulfill each instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into a single PDF file.
5. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your first and last name into the file name (e.g., "FirstLast_A02_CodingBasics.Rmd") prior to submission.



## Basics Day 1

1. Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name. 

2. Compute the mean and median of this sequence. 

3. Ask R to determine whether the mean is greater than the median. 

4. Insert comments in your code to describe what you are doing.

```{r}
#1. generate the sequence from 1 to 100, increasing by 4
seq(1, 100, 4)
sequence100_by4 <- seq(1, 100, 4)
sequence100_by4

#2. compute the mean and median of the above sequence
meanseq <- mean(sequence100_by4)
medianseq <- median(sequence100_by4)
meanseq
medianseq

#3. whether the above mean is greater than the above median
meanseq > medianseq

```

## Basics Day 2

5. Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6. Label each vector with a comment on what type of vector it is.

7. Combine each of the vectors into a data frame. Assign the data frame an informative name.

8. Label the columns of your data frame with informative titles.

```{r}

# 5 and 6
# 5-a creat 10 student names
StudentNameVect <- c ("Jack","Peter", "April", "John", "Rebacca","Julia","Max", "Lisa", "Grace", "Pari")
StudentNameVect
class(StudentNameVect) #6 the type of the vector of StudentNameVect

# 5-b create the vector of student test scores
StudentScoreVect <- ceiling(runif(10, min=0, max=100)) # get 10 random number from 0 to 100 and round up!
StudentScoreVect
class(StudentScoreVect) #6 the type of the vector of StudentScoreVect

# 5-c whether they have passed the test
IfPassVect <- StudentScoreVect > 50
IfPassVect
class(IfPassVect) #6 the type of the vector of IfPassVect

#7 get a data frame! #8 get beautiful names for columns!

#creat a dataframe
StudentTest <- data.frame ("StudentName" = StudentNameVect, "StudentScore" = StudentScoreVect, "IfPass" = IfPassVect) # to combine the data frame as a whole
StudentTest

is.data.frame(StudentTest) # to verify!
View(StudentTest) # to have a look!

```

9. QUESTION: How is this data frame different from a matrix?

> Answer: A data frame can contain different vector types, for example, a data frame could be a combination with colunmns with type of numeric, logical, or character, while a matrix can only contain one type of the vector (only nuneric or logical or other type at the same time)

10. Create a function with an if/else statement. Your function should determine whether a test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. Hint: Use `print`, not `return`. The name of your function should be informative.

11. Apply your function to the vector with test scores that you created in number 5. 

```{r}
IfPass_func <- function(x){
  ifelse(x>50, print("pass"), print("fail"))
} # 10 the function
Passorfail <- IfPass_func(StudentTest$StudentScore)
Passorfail # 11 Vector generation
```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: They all worked! We can choose either to enter "If x>50 {print ("pass")}, else{print("fail")}, or just use ifelse as I've used above.

Comments for the assignment: It is too much to finish all past videos and assignments in two days for a poor beginner (I just registered for the course before the drop/off ddl TvT), life is so hard!
